---
title: "Working notes"
output:
  github_document:
    html_preview: false
---

```{r setup}
library(icesTAF)
library(ggplot2)
library(mgcv)
taf.library(FLCore)
taf.library(FLa4a)
cat("buit on", date())
```

## Model building and simulating

Define a model for survey catchability - choose a range first that
only has ages for demonstration purposes.

```{r new-submodel}
qmod <-
  submodel(
    name = "trawl",
    range = c(min = 1, max = 10, minyear = 2017, maxyear = 2017),
    formula = ~ s(age, k = 4)
  )
```

we can visualise the various shapes this can take by simulating
from the 'prior' parameter space, which we are thinking of here
as multivariate normal with mean vector zero and identity matrix
variance covariance matrix. By setting the variance of the intercept
term to zero we can focus on the shape rather then the level of the
curve.  Note that the variance covariance is a 3D array, with the
3rd dimension potentially contianing simulated variance covariance
matrices, or variance covariance matrices arising from fits to
multiple simulated datasets.

```{r}
vcov(qmod)["(Intercept)", "(Intercept)", 1] <- 0
```

simuate from the object and visualise it

```{r fitted-values}
ggplot(
  as.data.frame(
    genFLQuant(qmod, nsim = 50, seed = 1323)
  ),
  aes(x = age, y = data, group = iter)
) +
  geom_line(colour = alpha("red", alpha = 0.3)) +
  labs(title = "Samples from the q model space (log scale with Intercept = 0)")
```

### Constraining q above a fixed age

in this example we use a formula that replaces ages above 8 with 8,
and therefore constrains q above 8 to be equal to q at age 8.

```{r new-submodel-flat}
qmod <-
  submodel(
    name = "trawl-trunc",
    range = c(min = 1, max = 10, minyear = 2017, maxyear = 2017),
    formula = ~ s(replace(age, age > 8, 8), k = 4)
  )
```

Again, by setting the variance of the intercept
term to zero we can focus on the shape rather then the level of the
curve.

```{r}
vcov(qmod)["(Intercept)", "(Intercept)", 1] <- 0
```

simuate from the object and visualise it

```{r fitted-values2}
ggplot(
  as.data.frame(
    genFLQuant(qmod, nsim = 50, seed = 1323)
  ),
  aes(x = age, y = data, group = iter)
) +
  geom_line(colour = alpha("red", alpha = 0.3)) +
  labs(title = "Samples from the q model space (log scale with Intercept = 0)")
```

an alternative to fixing the age in the formula is to supply a
truncated age as a covariate.  This is done by adding a data.frame
of covariates to the submodel, much in the same way that the `data`
argument in `lm()` provides covariate data.

To add covariates, in this case it is easier to create the data.frame
from the previous submodel, but it can be done in one

```{r new-submodel-with-covars}
qmod2 <-
  submodel(
    name = "trawl-trunc",
    range = c(min = 1, max = 10, minyear = 2017, maxyear = 2017),
    formula = ~ s(age_trunc, k = 4)
  )

data <- as.data.frame(qmod2, drop = TRUE)
data$age_trunc <- replace(data$age, data$age > 8, 8)

qmod2 <-
  submodel(
    qmod2,
    covariates = data
  )
vcov(qmod2)["(Intercept)", "(Intercept)", 1] <- 0


ggplot(
  as.data.frame(
    genFLQuant(qmod2, nsim = 50, seed = 1323)
  ),
  aes(x = age, y = data, group = iter)
) +
  geom_line(colour = alpha("red", alpha = 0.3)) +
  labs(title = "Samples from the q model space (log scale with Intercept = 0)")
```

### filling out over years

In order to create a q model that operates over several years, it is
just a case of extending the range of the submodel